\documentclass[12pt]{article}

\usepackage{amsmath, amsthm, amssymb,amscd}
\usepackage{setspace}
\usepackage{enumerate}
\usepackage[margin=1in]{geometry}


\newtheoremstyle{mystyle}% name
    {2ex}       % Space above
    {2ex}       % Space below
    {}          % Body font
    {}          % Indent amount (empty = no indent, \parindent = para indent)
    {\bfseries} % Thm head font
    {.}         % Punctuation after thm head
    {\newline}  % Space after thm head: \newline = linebreak
    {}          % Thm head spec

\theoremstyle{mystyle}
\newtheorem*{claim}{Claim}
\doublespace

% ------------------------------------------------------------------------------
% Begin Document
% ------------------------------------------------------------------------------
\title{Math 300 - Essay 4}
\author{George Lamperis}
\date{}

\begin{document}
\maketitle

Because of what the visitor says, the islanders will deduce that there is at
least one person on the island with blue eyes. First, suppose only one person on
the island has blue eyes. This person would look around at the other islanders
and notice that they all have brown eyes. Since this person knows that there is 
at least one person with blue eyes, this person will conclude that their eyes
must be blue. Then, at noon, this islander would leave.

Now suppose that there are two people who have blue eyes. Both will see each
other and notice that the other has blue eyes, while the rest of the islanders
have brown eyes. They both deduce that ``If I don't have blue eyes, then the
islander I see with blue eyes will leave today''. When this does not happen,
both will deduce that they have blue eyes, and both of them leave the following
day.

Let's examine one more case: suppose there are three islanders with blue eyes.
Each of these three will reason that, ``If I don't have blue eyes, then the
two with blue eyes will leave tomorrow.'' When nobody leaves on day 2, each of
them concludes that they have blue eyes, and all three leave the island on day
3.

\begin{claim}
  If there are $n$ islanders with blue eyes, then on the $n$-th day, all the
  blue-eyed islanders will realize that they have blue eyes. Thus, on the $n$-th
  day, all the islanders with blue eyes must leave.
\end{claim}
\begin{proof}
  If there is only one person with blue eyes, he will look at the other
  islanders and realize that none of them have blue eyes. Hence, this person
  concludes that he must have blue eyes, and thus leaves the island at noon.
  
  Now suppose that if there are $n > 1$ people on the island who have blue eyes,
  then on the $n$-th day, they all deduce their eye color is blue.
  
  Then if there are $n+1$ people with blue eyes, each will look around at the
  other islanders and notice $n$ of them with blue eyes.
  
  Each of these people deduces that either there are only $n$ people on the
  island with blue eyes, or that they have blue eyes too. So they all wait $n$
  days. When nobody with blue eyes has left the island, each of them now knows
  that they have blue eyes too. And so, on the $n+1$-th day, all $n+1$ blue-eyed
  islanders leaves.
\end{proof}


\end{document}
